---
title: "Lab Week 5: Time Series & A Map!"
author: "Keene Morrow"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach packages
```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US Energy Consumption for Renewables
```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>%
  janitor::clean_names()
```

- Make description all lowercase
- Only keep observations where `description` contains "consumption"
- Remove observation where `description` contains "total"

```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>%
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) # can also use negate = TRUE instead of !
```

### Convert `yyyymm` to date

```{r}
renew_date <- renew_clean %>%
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% #automatically adds day as the first of the month
  mutate(month_sep = yearmonth(yr_mo_day)) %>% # store a tsiblle specific class yearmonth in prep
  mutate(value = as.numeric(value)) %>%
  drop_na(month_sep, value)

# parse the year and month for use later.
renew_parsed <-renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))

```

### Make a ggplot

Make, then save, the baseline plot as `renew_gg`:

```{r}
renew_gg <- ggplot(data = renew_date,
                   aes(x = month_sep,
                       y = value,
                       group = description)) + # redundant due to aes(color = description) in next line
  geom_line(aes(color = description)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 350))
  

renew_gg
```

### Updating colors with `paletteer` Palettes

The above data is discrete, so we're looking for discrete palettes in paletteer using `view(palettes_d_names)`

```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::figmtn")

renew_gg +
  scale_color_paletteer_d("palettesForR::Inkscape")

renew_gg +
  scale_color_paletteer_d("trekcolors::klingon")
	
renew_gg +
  scale_color_paletteer_d("LaCroixColoR::Tangerine")
```

### Coerce renew_parsed to tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at time series data in a couple way
```{r}
renew_ts %>% autoplot(value) # if key is indicated above, auto groups data
renew_ts %>% gg_subseries(value)
# renew_ts %>% gg_season(value) # broken ;/

ggplot(data = renew_parsed,
       aes(x = month,
           y = value,
           group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~ description,
             ncol = 1,
             scales = "free",
             strip.position = "right")

```

### Hydroelectric Power Only
```{r}
hydro_ts <- renew_ts %>%
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(hydro_ts,
       aes(x = month,
           y = value,
           group = year)) +
  geom_line(aes(color = year))
```

### Quarterly Average Consumption of Hydroelectric Power
```{r}
hydro_quarterly <- hydro_ts %>%
  index_by(year_qu = ~(yearquarter(.))) %>% #coerce to quarters based on existing key (.)
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose Hydro Time Series data

```{r}
dcmp <- hydro_ts %>%
  model(STL(value ~ season(window = 5)))

# show components from decomposition and plot them
components(dcmp) %>% autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

Autocorellation Function (ACF)
```{r}
hydro_ts %>%
  ACF(value) %>%  # the function from feast
  autoplot()
```
Shows high seasonal correlation, lag of 12 has highest corr

# DANGER ZONE
More research required for future endeavours
Be sure to consider how valid predictions may be
Comparisons of model types can  reinforce predictions powerfully

```{r}
hydro_model <- hydro_ts %>%
  model(
    ARIMA(value)
  ) %>%
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot # plot just model
hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010)) # include data used to make predictions
```


### Make a world map!
using mapview
```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

